SMT-Based Generation of Symbolic Automata \thanks{This work was partially funded by the Associated Team FM4CPS between INRIA and ECNU, Shanghai}

### 2019-03-30 2017-Xudong-SMT/ActaInformatica.tex
Acta Inf. paper (minor): removed the mention of priorities)
preuve Th2

\newcommand{\noteSB}[2][color=green!40, size=\tiny]{\todo[#1]{{\bf Simon: } {#2}}}
\newcommand{\noteEM}[2][color=blue!40, size=\tiny]{\todo[#1]{{\bf Eric: } {#2}}}
\newcommand{\noteInSB}[2][inline,color=green!40]{\todo[#1]{{\bf Simon: } {#2}}}
\newcommand{\noteInEM}[2][inline,color=green!40]{\todo[#1]{{\bf Eric: } {#2}}}
\newcommand{\todoMargin}[2][color=red!60, size=\tiny]{\todo[#1]{{\bf
      ToDo: } {#2}}}
\newcommand{\hlEM}[1]{{\sethlcolor{green}\hl{#1}}}

@@

\input{ActaXFIG/macros.tex}         % Definitions of macros used in XFig figures
\usepackage{macrospNets}

@@

\begin{example}[Open transitions]
  Transition $OT_1$ of the pNet $EnableStateCompLeft$ is generated by
  combining the first vector of its root node with
  the second vector of its second node. Its generated predicate is
  trivially true. The global states show that controler $CState1$ is
  changing from state 0 to 1.
  \label{OT:example-1}
        \begin{mathpar}
	  OT_1  =
          \inferrule*[myfraction=\reddottedrule]
	{\{\xrightarrow{\delta(x2)}_P ~~
          \xrightarrow{acc(x2)}_Q \} ~~
          [\forall z. acc(x2) \neq \delta(y1)]
        }
	{\ostate{0,0} \xrightarrow {\underline{\delta(x2)}} \ostate{1,0}}
	\end{mathpar}

@@

  \label{OT:example-2}
        \begin{mathpar}
	  OT_2  =
          \inferrule*[myfraction=\reddottedrule]
	{\{\xrightarrow{\delta(x2)}_P ~~
          \xrightarrow{acc(x2)}_Q \} ~~
          [v_1=0 \land v_2 = 0 \land \forall z. acc(x2) \neq
            \delta(z)] ~~ \{v1:=1\}
        }
	{\ostate{0,0} \xrightarrow {\underline{\delta(x2)}} \ostate{0,0}}
	\end{mathpar}
        where $v_1$ and $v_2$ are respectively the state variables of
        the first and second EnableData pNets. Indeed the global
        state space of this version is reduced to $\ostate{0,0}$, and
        the state change is replaced by the assignment ``$v_1:=1$''.

@@

\begin{algorithm}[h]
  \caption{Open Automaton Generation}
  \label{alg1}
\begin{algorithmic}[1]
\Require A pNet $P$ (can be a PLTS, but not a hole)
        %\qquad \hl{\textbackslash\textbackslash~SB: was `` pn ''}
\State Initialize sets $U=\{\mathit{InitState}(P)\}$ and $E=\emptyset$,
for unexplored and explored global states, respectively; $L=\emptyset$ for the resulting OTs;
\While{!$\mathit{isEmpty}(U)$}
	\State Choose $S$ in $U$; remove $S$ from $U$, add $S$ to $E$;
        %\qquad \hl{\textbackslash\textbackslash~SB: was `` Choose $S$ in $U$ ''}
	\State $\mathit{OTs} = \mathit{MakeTransitions}(P, S)$;
	\For {each $\mathit{OT} \in \mathit{OTs}$}
          \State Add $\mathit{OT}$ to $L$;
          \State Let $S'$ be the target state of $\mathit{OT}$
          Add $S'$ to the unexplored states if necessary:
	  \IfThen{($S' \not\in U \cup E$)}{Add $S'$ into $U$;}
	  \EndFor
\EndWhile
\State \QIN{$\mathit{L1}$ = \texttt{filterUnsatTransitions}($\mathit{L,Assignements}$)}
\State \QIN{$\mathit{OA1}$ = \texttt{makeReachableSubAutomaton}($\mathit{InitState}(P)$,$\mathit{L1}$)}
\State \Return $\mathit{OA1}$;

@@

\ERIC{Before the SAT checking step, it may contain symbolic transitions that
do not have valid ground instanciations. This is not logically
incorrect, but is naturaly not an optimal representation. So we prune
these UNSAT transitions: at the end of the open automaton generation,
the predicate of each OT is translated 
into SMTlib assertions (see Section~\ref{section:pruning}), and checked for
satisfiability (function \texttt{filterUnsatTransitions}). 
Then we compute the set of reachable transitions and states to build
the final open automaton.}

@@

%\st{Each of these possible filters is more precise than the previous
%  ones. But e}
Even when unification succeeds, it is still possible that
the resulting predicate would be unsatisfiable, because of some
incompatibility involving the guards. In our algorithm, we choose to
apply only the simplest filter inside the \emph{Matching} method
(before applying the predicate and OT construction). Matching and
unification  will be checked later, together with the guards collected
from synchronisation vectors and from PLTS transtions, using the
satisfiability check in the SMT engine.

@@

For a global state in the open automaton, the set of state
variables (which may be used in an outgoing transition) 
is the disjoint union
of the sets of state variables of the individual PLTS states constituting this
global state. \ERIC{The final assignment sets are assembled when building the
open automaton from the reachable transitions, see below.}

@@

Function \texttt{filterUnsatTransitions}, in line 11 of
Algorithm\ref{alg1} makes use of the Z3 engine to implement the
satisfiability checking, for all the OTs computed that far, and
returns the set $L1$ of satisfiable OTs.

\subsection{Building the Open Automaton}

\ERIC{The set $L1$ may contain unreachable transitions. To remove those transitions, the last step of the
algorithm builds the final open automaton via a traversal of $L1$,
starting at the initial state $InitState(p)$ as defined in
page~\pageref{def-states}. More precisely, this is a residual
algorithm building a graph from the reachable transitions in $L1$. 
%and states are the set of target states of the transitions. 
Each state in the graph is decorated with a set of its local
variables, and a list of their possible assignments,
gathered during the traversal using the ``Post'' parts of the
OTs.}

\ERIC{A further reduction may now be gained by using the information from
these sets of possible assignments: adding this information when
checking satisfiability of transitions leaving a given state may
result in more Unsat transitions. The set of possible
assignments computed by the algorithm is an over approximation of
possible values of the 
variables of such state, so narrowing it means we get a better
approximation, and potentially more information, which can be used in
the SAT checking of subsequent transitions. In turn, this may reduce the possible
assignments of the corresponding target states. In theory, we should
iterate this procedure, each iteration reducing the set of
transitions and the sets of assignments; the final open automaton
would be the fixed point of this iteration. In practice, this would be
expensive, and it is not incorrect to leave some Unsat OTs in the
result. From a  practical point of view, very little will be gained to compute the
fixed point, so we only perform one step of iteration. Further algorithms 
(e.g. model-checking or equivalence checking) will be able to use the
full information from the predicates and the assignment sets.}

@@

But there is more to this. Due to the symbolic nature of the
  transitions, reachability is not as easy to define as in standard
  models, as illustrated by the definition above. By avoiding building
  the OTs coming from some unreachable states, we also avoid adding
  the corresponding assignments to the variables of their target states.
But we \emph{cannot} use the
assignment information in the ``on-the-fly'' SAT checking, because
they cannot be complete before exploring (at least) all the reachable
state space. We first need to compute a first over-approximation of
the reachable state space, and assignment sets; this corresponds to
the set of OTs $L2$ in Algorithm~\ref{alg3}, line 15.
So the smart algorithm proceeds in two steps: during the
first step, the SAT checking is done without using any assignment set
information. Then after the completion of this step, we perform another
SAT checking of all transitions from step1, using the
assignment information, in the same way it is done at the end of
Algorithm~\ref{alg1}.

There is one case much simpler, that is when the pNets contains no
state variable assignment. Then the set of transitions computed
on-the-fly is already guaranted to be the final reachable set (line 13).

\begin{algorithm}[h]
  \caption{``Smart'' Reachable Open Automaton Generation}
  \label{alg3}
\begin{algorithmic}[1]
\Require A pNet $P$ (cannot be a hole)
        %\qquad \hl{\textbackslash\textbackslash~SB: was `` pn ''}
\State Initialize sets $U=\{\mathit{InitState}(P)\}$ and $E=\emptyset$,
for unexplored and explored global states, respectively; $L=\emptyset$ for the resulting OTs;
\While{!$\mathit{isEmpty}(U)$}
	\State Choose $S$ in $U$; remove $S$ from $U$, add $S$ to $E$;
        %\qquad \hl{\textbackslash\textbackslash~SB: was `` Choose $S$ in $U$ ''}
	\State $\mathit{OTs} = \mathit{MakeTransitions}(P, S)$;
%%        \State Store $\mathit{OT}$ in $L$; 
	\For {each $\mathit{OT} \in \mathit{OTs}$}
        Check satisfiability of $\mathit{OT}$ using the SMT solver;
        \IfThen {$SAT(\mathit{OT})$}
                {\State \{Add $\mathit{OT}$ to $L$;
                  \State Let $S'$ be the target global state of $\mathit{OT}$
	          \IfThen{($S' \not\in U \cup E$)}{Add $S'$ into $U$;}
                  \}}
	\EndFor
\EndWhile
%%\State Prune the unsatisfiable open transitions in $L$ using the SMT solver;
\IfThenElse {(NonEmptyAssignmentSet($\mathit{L}$))}
        {\State \quad\ \Return $\mathit{OA2}$=makeAutomaton($\mathit{InitState}(P),L)$;}
        {\State \quad\ $\mathit{L2}$=filterUnsatTransitions($\mathit{L,Assignments}$);
         \State \quad\ \Return $\mathit{OA2}$=makeReachableSubAutomaton($\mathit{InitState}(P),L2)$};
\end{algorithmic}  
\end{algorithm}

@@

The easy case is when the system does not contain assignments to state
variables. This does not mean ``data-independency'': we can still have
guards in the pLTS transitions and in the synchronisation vectors,
making the behaviour depending on the value of input variables. But
the open automaton needs not to collect the set of possible values dealing
with the values of local variables, so we can directly compute the
reachable part of the automaton using the on-the-fly algorithm.

@@

 Let $pn=$ be a pNet. The set of states 
 managed by the two algorithms both are subsets of the cartesian
 product of the pLTS states, as defined in page \pageref{def-states}. In a
 similar way, the set of state variables attached to each state are
 also equal. Comparing  
 transitions requires a little more care, because within the
 MakeTransitions function, fresh variables are created everytime a
 synchronisation vector is used (this occurs within the $MkPred$ and
 $matching$ functions). However, given a global state, the call of
 $MakeTransitions$ will always returns exactly the same set of OTs, up to
renaming of these fresh variables

  Let us first define some notations.
 \begin{itemize}
    \item $L1$: the original set of transitions by Algorithm 1
      (i.e. before line 11), 
    \item $L1_{sat}$:  the original set of transitions by Algorithm 1, after
      satisfiability check (i.e. after line 11),
    \item $L1_{reach}$: the final set of OTs of Algorithm 1, as computed
      by line 12.
    \item $L2_{sat}$: the original set of transitions by Algorithm 3
      (i.e. before line 13), 
    \item $L2_{reach}$: the final set of OTs of Algorithm 2, as computed
          by line 13.
\end{itemize}

\paragraph{States:} for each of these, define $States: \langle OT\rangle\rightarrow\langle State\rangle$ which is the function computing the set of target states in a
set of OTs.

Between two transition sets of the algorithms, define
$Rename:  L1_{reach} \rightarrow L2_{reach}$ which is the function
renaming fresh variables.

When the pNets contains no assignment, there is no need for the last
filtering step, so $L1_{sat}=L1_{reach}$ and $L2_{sat}=L2_{reach}$.

According to the definition of global states in
page~\pageref{def-states}, the global state
$\ostate{\tilde{s}}=\ostate{\tilde{s_0...s_n}}$ is a composition of $n$
pLTS states $s_{i}^{i\in [0,n]}$.

From the structure of the main loop of the algorithms, we can get that algorithm~\ref{alg1} computes less transitions than algorithm~\ref{alg3}, i.e. $L2_{reach}\subseteq  L1$, and as a consequence $States(L2_{reach})\subseteq  States(L1)$, so we have:
\[\forall \ostate{\tilde{u}}\in States(L2_{reach}), \exists\ostate{\tilde{t}}\in States(L1), \ostate{\tilde{t}}=\ostate{\tilde{a_0...a_n}}=\ostate{\tilde{u}}.\]

The reachability function $computeReachableSubAutomaton$ being called
with the same initial state $\ostate{\tilde{t_0}}$, this state is the
initial state of both $L1_{reach}$ and $L2_{reach}$.

%  If there is an initial state $\ostate{\tilde{t_0}}\in States(L2_{reach})$ then the initial state of the original automaton $\ostate{\tilde{t_0}}\in States(L1_{reach})$ must exists.

\noteInEM{What you are doing below is a proof by recursion on the
  length of the path reaching a given OT, isn't it ? I suggest you
  explicite this and conclude $L2_{reach}\subseteq  L1_{reach}$.}

Suppose that  $\ostate{\tilde{t}}\in States(L2_{reach})$. It is reachable in the result automaton of Algorithm 3, so there exists a sequence of transitions 
\[\ostate{\tilde{t_0}}\xrightarrow{l_1}\ostate{\tilde{t_1}}\xrightarrow{l_2}...\xrightarrow{l_{n}}\ostate{\tilde{t}}\] where in each label of the transition $Pred_{l_i}^{i\in [1,n]}$ is satisfiable. For the same state $\ostate{\tilde{t}}\in States(L1_{reach})$, there exists a sequence of transitions \[\ostate{\tilde{t_0}}\xrightarrow{l'_1}\ostate{\tilde{t_1}}\xrightarrow{l'_2}...\xrightarrow{l'_{n}}\ostate{\tilde{t}}\]where for each transition $Rename(\ostate{\tilde{t_{i-1}}}\xrightarrow{l_{i}}\ostate{\tilde{t_{i}}}) = \ostate{\tilde{t_{i-1}}}\xrightarrow{l'_{i}}\ostate{\tilde{t_{i}}}$ and $Pred_{l'_i}^{i\in [1,n]}$ is satisfiable. 
From the structure of the algorithms, 
\[\forall \ostate{\tilde{t}}\in States(L2_{reach}), \ostate{\tilde{t}}\xrightarrow{l'_{n+1}}\ostate{\tilde{t'}}\in L2_{reach}\]
\[\exists \ostate{\tilde{t}}\xrightarrow{l_{n+1}}\ostate{\tilde{t'}}\in L1 \quad s.t.\quad Rename(\ostate{\tilde{t}}\xrightarrow{l_{n+1}}\ostate{\tilde{t'}}) = \ostate{\tilde{t}}\xrightarrow{l'_{n+1}}\ostate{\tilde{t'}})\]
where $Pred_{l_{n+1}}$ and $Pred_{l'_{n+1}}$ are satisfiable.

Then we can get a new longer sequence \[\ostate{\tilde{t_0}}\xrightarrow{l_1}\ostate{\tilde{t_1}}\xrightarrow{l_2}...\xrightarrow{l_{n}}\ostate{\tilde{t}}\xrightarrow{l_{n+1}}\ostate{\tilde{t}}\] \[\Rightarrow\ostate{\tilde{t'}}\in States(L1_{reach})\wedge\ostate{\tilde{t}}\xrightarrow{l_{n+1}}\ostate{\tilde{t'}}\in L1_{reach}\in L1.\] So for arbitrary transition in $L2_{reach}$, there is a same transition in $L1_{reach}$ after renaming its label, then we get
$L2_{reach}\subseteq  L1_{reach}$.

Conversely,
\[\forall \ostate{\tilde{t'}}\notin States(L2_{reach}), \ostate{\tilde{t}}\xrightarrow{l'}\ostate{\tilde{t'}}\notin L2_{reach},\quad\exists \ostate{\tilde{t'}}\in States(L1), \ostate{\tilde{t}}\xrightarrow{l}\ostate{\tilde{t'}}\in L1\]
where $Pred_{l}$ and $Pred_{l'}$ are not satisfiable. As pruning function in Algorithm 1 will remove all the unsatisfiable transitions, then $\ostate{\tilde{t}}\xrightarrow{l}\ostate{\tilde{t'}}\notin L1_{reach}$.

>>>>

\newcommand{\noteSB}[2][color=green!40, size=\tiny]{\todo[#1]{{\bf
      Note: } {#2}}}
\newcommand{\noteEM}[2][color=blue!40, size=\tiny]{\todo[#1]{{\bf Eric: } {#2}}}
\newcommand{\noteInSB}[2][inline,color=green!40]{\todo[#1]{{\bf Simon: } {#2}}}
\newcommand{\noteInEM}[2][inline,color=green!40]{\todo[#1]{{\bf Eric: } {#2}}}
\newcommand{\todoMargin}[2][color=red!60, size=\tiny]{\todo[#1]{{\bf
      ToDo: } {#2}}}
\newcommand{\hlEM}[1]{{\sethlcolor{green}\hl{#1}}}

@@

\input{XFIG/macros.tex}         % Definitions of macros used in XFig figures
\usepackage{macrospNets}

@@

\begin{example}[Open transitions]
  Transition $OT_2$ of the pNet $EnableStateCompLeft$ is generated by
  combining the first vector of its root node with
  the second vector of its second node. The global states show that
  controler $CState1$ is 
  changing from state 0 to 1.
  \label{OT:example-1}
        \begin{mathpar}
	  OT_2  =
          \inferrule*[myfraction=\reddottedrule]
	{\{\xrightarrow{\delta(x2)}_P ~~
          \xrightarrow{acc(x2)}_Q \} ~~
          [\forall z. acc(x2) \neq \delta(y1)]
        }
	{\ostate{0,0} \xrightarrow {\underline{\delta(x2)}} \ostate{1,0}}
	\end{mathpar}

@@

  \label{OT:example-2}
        \begin{mathpar}
	  OT_2'  =
          \inferrule*[myfraction=\reddottedrule]
	{\{\xrightarrow{\delta(x2)}_P ~~
          \xrightarrow{acc(x2)}_Q \} ~~
          [v_1=0 \land v_2 = 0 \land \forall z. acc(x2) \neq
            \delta(z)] ~~ \{v1:=1\}
        }
	{\ostate{0,0} \xrightarrow {\underline{\delta(x2)}} \ostate{0,0}}
	\end{mathpar}
        where $v_1$ and $v_2$ are respectively the state variables of
        the first and second EnableData pNets. Indeed the global
        state space of this version is reduced to $\ostate{0,0}$, and
        the state change is replaced by the assignment ``$v_1:=1$''.

@@

\begin{algorithm}[h]
  \caption{Open Automaton Generation}
  \label{alg1}
\begin{algorithmic}[1]
\Require A pNet $P$ (can be a PLTS, but not a hole)
        %\qquad \hl{\textbackslash\textbackslash~SB: was `` pn ''}
\State Initialize sets $U=\{\mathit{InitState}(P)\}$ and $E=\emptyset$,
for unexplored and explored global states, respectively; $L=\emptyset$ for the resulting OTs;
\While{!$\mathit{isEmpty}(U)$}
	\State Choose $S$ in $U$; remove $S$ from $U$, add $S$ to $E$;
        %\qquad \hl{\textbackslash\textbackslash~SB: was `` Choose $S$ in $U$ ''}
	\State $\mathit{OTs} = \mathit{MakeTransitions}(P, S)$;
	\For {each $\mathit{OT} \in \mathit{OTs}$}
          \State Add $\mathit{OT}$ to $L$;
          \State Let $S'$ be the target state of $\mathit{OT}$
          Add $S'$ to the unexplored states if necessary:
	  \IfThen{($S' \not\in U \cup E$)}{Add $S'$ into $U$;}
	  \EndFor
\EndWhile
\State $\mathit{L1}$=filterUnsatTransitions($\mathit{L,Assignments}$);
\State $\mathit{OA_0}$=makeReachableSubAutomaton($\mathit{InitState}(P),L1)$;
\Repeat $\mathit{OA_n}$=RefineReachableSubAutomaton($\mathit{AO_{n-1}}$)
\Until  $\mathit{OA_n}=\mathit{OA_{n-1}}$; \Comment Fixpoint reached
\State \Return $\mathit{OA_n}$;

@@

Before the SAT checking step, it may contain symbolic transitions that
do not have valid ground instanciations. This is not logically
incorrect, but is naturaly not an optimal representation. So we prune
these UNSAT transitions: at the end of the open automaton generation,
the predicate of each OT is translated 
into SMTlib assertions (see Section~\ref{section:pruning}), and checked for
satisfiability (function \texttt{filterUnsatTransitions}). 
Then we compute the set of reachable transitions and states to build
the final open automaton (more details in section~\ref{section:BuildingTheOpenAutomaton}).

@@

Even when unification succeeds, it is still possible that
the resulting predicate would be unsatisfiable, because of some
incompatibility involving the guards. In our algorithm, we choose to
apply only the simplest filter inside the \emph{Matching} method
(before applying the predicate and OT construction). Matching and
unification  will be checked later, together with the guards collected
from synchronisation vectors and from PLTS transtions, using the
satisfiability check in the SMT engine.

@@

For a global state in the open automaton, the set of state
variables (which may be used in an outgoing transition) 
is the disjoint union
of the sets of state variables of the individual PLTS states constituting this
global state. The final assignment sets are assembled when building the
open automaton from the reachable transitions, see below.

@@

Function \texttt{filterUnsatTransitions}, in line 11 of
Algorithm\ref{alg1}, and function \\
\texttt{RefineReachableSubAutomaton}
in line 13, make use of the Z3 engine to implement the
satisfiability checking, for all the OTs computed that far.

\subsection{Building the Open Automaton}
\label{section:BuildingTheOpenAutomaton}

The set $L1$ may contain unreachable transitions. To remove those transitions, the last step of the
algorithm builds the final open automaton via a traversal of $L1$,
starting at the initial state $InitState(p)$ as defined in
page~\pageref{def-states}. More precisely, this is a residual
algorithm building a graph from the reachable transitions in $L1$. 
%and states are the set of target states of the transitions. 
Each state in the graph is decorated with a set of its local
variables, and a list of their possible assignments,
gathered during the traversal using the ``Post'' parts of the
OTs.

A further reduction may now be gained by using the information from
these sets of possible assignments: adding this information when
checking satisfiability of transitions leaving a given state may
result in more Unsat transitions. The set of possible
assignments computed by the algorithm is an over approximation of
possible values of the 
variables of such state, so narrowing it means we get a better
approximation, and potentially more information, which can be used in
the SAT checking of subsequent transitions. In turn, this may reduce the possible
assignments of the corresponding target states. In
algorithm~\ref{alg1}, we iterate this procedure, each iteration
reducing the set of 
transitions and the sets of assignments; the final open automaton
is the fixed point of this iteration.

In practice, this could be
expensive, and it is not incorrect to leave some Unsat OTs in the
result. From a  practical point of view, very little will be gained to compute the
fixed point, so we would only perform one step of iteration. Further algorithms 
(e.g. model-checking or equivalence checking) will be able to use the
full information from the predicates and the assignment sets.

@@

But there is more to this. Due to the symbolic nature of the
  transitions, reachability is not as easy to define as in standard
  models, as illustrated by the definition above. By avoiding building
  the OTs coming from some unreachable states, we also avoid adding
  the corresponding assignments to the variables of their target states.
But we \emph{cannot} use the
assignment information in the ``on-the-fly'' SAT checking, because
they cannot be complete before exploring (at least) all the reachable
state space. We first need to compute a first over-approximation of
the reachable state space, and assignment sets; this corresponds to
the automaton $OA_0$ in Algorithm~\ref{alg3}, line 15.
So the smart algorithm proceeds in two steps: during the
first step, the SAT checking is done without using any assignment set
information. Then after the completion of this step, we refine the
reachability computation, using the
assignment information as described in the previous section.

\begin{algorithm}[h]
  \caption{``Smart'' Reachable Open Automaton Generation}
  \label{alg3}
\begin{algorithmic}[1]
\Require A pNet $P$ (cannot be a hole)
        %\qquad \hl{\textbackslash\textbackslash~SB: was `` pn ''}
\State Initialize sets $U=\{\mathit{InitState}(P)\}$ and $E=\emptyset$,
for unexplored and explored global states, respectively; $L=\emptyset$ for the resulting OTs;
\While{!$\mathit{isEmpty}(U)$}
	\State Choose $S$ in $U$; remove $S$ from $U$, add $S$ to $E$;
        %\qquad \hl{\textbackslash\textbackslash~SB: was `` Choose $S$ in $U$ ''}
	\State $\mathit{OTs} = \mathit{MakeTransitions}(P, S)$;
	\For {each $\mathit{OT} \in \mathit{OTs}$}
        Check satisfiability of $\mathit{OT}$ using the SMT solver;
        \IfThen {$SAT(\mathit{OT})$}
                {\State \{Add $\mathit{OT}$ to $L$;
                  \State Let $S'$ be the target global state of $\mathit{OT}$
	          \IfThen{($S' \not\in U \cup E$)}{Add $S'$ into $U$;}
                  \}}
	\EndFor
\EndWhile
\State $\mathit{L2}$=filterUnsatTransitions($\mathit{L,Assignments}$);
\State $\mathit{OA_0}$=makeReachableSubAutomaton($\mathit{InitState}(P),L2)$;
\Repeat $\mathit{OA_n}$=RefineReachableSubAutomaton($\mathit{AO_{n-1}}$)
\Until  $\mathit{OA_n}=\mathit{OA_{n-1}}$;
\State \Return $\mathit{OA_n}$;
%\EndIf
\end{algorithmic}  
\end{algorithm}

@@

The easy case is when the system does not contain assignments to state
variables. This does not mean ``data-independency'': we can still have
guards in the pLTS transitions and in the synchronisation vectors,
making the behaviour depending on the value of input variables. But
the open automaton needs not to collect the set of possible values dealing
with the values of local variables, so we can directly compute the
reachable part of the automaton using the on-the-fly algorithm (line 13).

@@

 Let $pn=$ be a pNet. The set of states 
 managed by the two algorithms both are subsets of the cartesian
 product of the pLTS states, as defined in page \pageref{def-states}. In a
 similar way, the set of state variables attached to each state are
 also equal. Comparing  
 transitions requires a little more care, because within the
 MakeTransitions function, fresh variables are created everytime a
 synchronisation vector is used (this occurs within the $MkPred$ and
 $matching$ functions). However, given a global state, the call of
 $MakeTransitions$ will always returns exactly the same set of OTs, up to
renaming of these fresh variables. When this is not ambiguous, we will
speak of equality (of transitions, and of open automata) when we
should say ``up to renaming of fresh variables''.

  Let us first define some notations.
 \begin{itemize}
    \item $L0$: the original set of transitions by Algorithm 1
      (i.e. before line 11), 
    \item $L1_{sat}$:  the original set of transitions by Algorithm 1, after
      satisfiability check (i.e. after line 11),
    \item $L1_{reach}$: the final set of OTs of Algorithm 1, as computed
      by line 12 (as $AO_0$). Without assignments the following
      refinement step will do nothing, the fixpoint is already reached.
    \item $L2_{sat}$: the original set of transitions by Algorithm 3
      (i.e. before line 13), 
    \item $L2_{reach}$: the final set of OTs of Algorithm 2, as computed
          by line 13. Here also $AO_n = AO_0$.
\end{itemize}

\paragraph{States:} for each of these, define
 $States: \overline{OT}\rightarrow\overline{State}$ the function
computing the set of target states of a set of OTs. \\
\noindent
\emph{Transitions:} between two sets of open transitions computed
by the algorithms, define $Rename:  L1_{reach} \rightarrow L2_{reach}$
which is the function renaming fresh variables.

When the pNets contains no assignment, there is no need for the last
filtering step, so $L1_{sat}=L1_{reach}$ and $L2_{sat}=L2_{reach}$.
We want to prove that $L1_{reach}=L2_{reach}$, that will imply
the equality of the 2 open automata.

\medskip
Let $S=\ostate{\overline{s}}=\ostate{\{s_{i}\}^{i\in Leaves(pn)}}$
range over the global states $States(L0)$, 
The only difference between the main loops of the algorithms is that 
algorithm~\ref{alg3} only retains the satisfiable OTs (line 6-7), so 
algorithm~\ref{alg3} computes less transitions than
algorithm~\ref{alg1}. Formaly $L2_{reach}\subseteq  L2_{sat} \subseteq L0$, and as a consequence $States(L2_{reach})\subseteq  States(L0)$, so we have:\\
\centerline{$\forall \ostate{\overline{u}}\in States(L2_{reach}), \exists\ostate{\overline{t}}\in States(L0), \ostate{\overline{t}}=\ostate{\overline{u}}$}

For a global state $S$, and a set of OTs $L$, define its distance
$|S|_L$ as the length $n$ of (one of the) shortest chain
$S_0\xrightarrow{l_1}S_1\xrightarrow{l_2}...S_{n-1}\xrightarrow{l_{n}}S$
  from the initial state $S_0$, with $\forall k\in [1..n], l_k \in L$.

\medskip
We will prove by induction on the distance from the original state,
that $\forall S \in L0. S \in States(L2_{reach}) \land |S|_{l2_{reach}}=n => S
\in States(L1_{reach})$, and vice-versa.

  The initial property is trivial, for $n=0$, both state sets of
  distance zero are reduced to $S_0$.

  Suppose our property is true for distance $n$, and
  suppose $|S|_{L2_{reach}}=n+1$, then there exist a chain
  $S_0\xrightarrow{l_1}S_1\xrightarrow{l_2}...S_{n}\xrightarrow{l_{n+1}}S$
  with all $l_k \in L2_{reach}$. In particular, as
  $L2_{reach}\subseteq L2_{sat}$, $l_{n+1} \in L2_{sat}$, meaning it
  it's predicate is satisfiable.

  By induction hypothesis, we get $S_{n} \in States(L1_{reach})$, so during the
  call of $makeReachableSubAutomaton$ in algorithm~\ref{alg1}, line 12, the OT
  $S_n \xrightarrow{l_{n}}S$ will be check and found SAT, so
  $S \in States(L1_{reach})$.

  Conversely if $|S|_{L1_{reach}}=n+1$, we get a similar chain, in
  $L1_{reach}$, with $S_n \xrightarrow{l_{n}}S, S_n \in
  States(L1_{reach})$, and $l_n$ satisfiable. By induction, we have
  $S_n \in States(L2_{reach})=States(L2_{sat})$, so $l_n$ will have
  been checked ``on-the-fly'', in line 6 of algorithm~\ref{alg3}, and
  the OT $S_n \xrightarrow{l_{n}}S$ added to $L2_{sat})$, and we
  conclude $S \in States(L2_{reach})$.

Remember that $L0$ is finite, so we conclude $L1_{reach}=L2_{reach}$, and $States(L1_{reach})=States(L2_{reach})$.

%  If there is an initial state $\ostate{\overline{t_0}}\in States(L2_{reach})$ then the initial state of the original automaton $\ostate{\overline{t_0}}\in States(L1_{reach})$ must exists.

%% \noteInEM{What you are doing below is a proof by recursion on the
%%   length of the path reaching a given OT, isn't it ? I suggest you
%%   explicite this and conclude $L2_{reach}\subseteq  L1_{reach}$.}

%% Suppose that  $\ostate{\overline{t}}\in States(L2_{reach})$. It is reachable in the result automaton of Algorithm 3, so there exists a sequence of transitions 
%% \[\ostate{\overline{t_0}}\xrightarrow{l_1}\ostate{\overline{t_1}}\xrightarrow{l_2}...\xrightarrow{l_{n}}\ostate{\overline{t}}\] where in each label of the transition $Pred_{l_i}^{i\in [1,n]}$ is satisfiable. For the same state $\ostate{\overline{t}}\in States(L1_{reach})$, there exists a sequence of transitions \[\ostate{\overline{t_0}}\xrightarrow{l'_1}\ostate{\overline{t_1}}\xrightarrow{l'_2}...\xrightarrow{l'_{n}}\ostate{\overline{t}}\]where for each transition $Rename(\ostate{\overline{t_{i-1}}}\xrightarrow{l_{i}}\ostate{\overline{t_{i}}}) = \ostate{\overline{t_{i-1}}}\xrightarrow{l'_{i}}\ostate{\overline{t_{i}}}$ and $Pred_{l'_i}^{i\in [1,n]}$ is satisfiable. 
%% From the structure of the algorithms, 
%% \[\forall \ostate{\overline{t}}\in States(L2_{reach}), \ostate{\overline{t}}\xrightarrow{l'_{n+1}}\ostate{\overline{t'}}\in L2_{reach}\]
%% \[\exists \ostate{\overline{t}}\xrightarrow{l_{n+1}}\ostate{\overline{t'}}\in L1 \quad s.t.\quad Rename(\ostate{\overline{t}}\xrightarrow{l_{n+1}}\ostate{\overline{t'}}) = \ostate{\overline{t}}\xrightarrow{l'_{n+1}}\ostate{\overline{t'}})\]
%% where $Pred_{l_{n+1}}$ and $Pred_{l'_{n+1}}$ are satisfiable.

%% Then we can get a new longer sequence \[\ostate{\overline{t_0}}\xrightarrow{l_1}\ostate{\overline{t_1}}\xrightarrow{l_2}...\xrightarrow{l_{n}}\ostate{\overline{t}}\xrightarrow{l_{n+1}}\ostate{\overline{t}}\] \[\Rightarrow\ostate{\overline{t'}}\in States(L1_{reach})\wedge\ostate{\overline{t}}\xrightarrow{l_{n+1}}\ostate{\overline{t'}}\in L1_{reach}\in L1.\] So for arbitrary transition in $L2_{reach}$, there is a same transition in $L1_{reach}$ after renaming its label, then we get
%% $L2_{reach}\subseteq  L1_{reach}$.

%% Conversely,
%% \[\forall \ostate{\overline{t'}}\notin States(L2_{reach}), \ostate{\overline{t}}\xrightarrow{l'}\ostate{\overline{t'}}\notin L2_{reach},\quad\exists \ostate{\overline{t'}}\in States(L1), \ostate{\overline{t}}\xrightarrow{l}\ostate{\overline{t'}}\in L1\]
%% where $Pred_{l}$ and $Pred_{l'}$ are not satisfiable. As pruning function in Algorithm 1 will remove all the unsatisfiable transitions, then $\ostate{\overline{t}}\xrightarrow{l}\ostate{\overline{t'}}\notin L1_{reach}$.
